# 6. Analysis of longitudinal data

We will use the two datasets that we developed in the meet_and_repeat R script.  
First, we will start with the dataset RATS. 

## RATS

```{r}
#We load the needed libraries
library(ggplot2)
library(tidyr)
library(dplyr)
ratsl <- read.csv(file = "~/IODS-project/data/ratsl.csv", row.names = 1)

#We convert the categorical variables into factors again
ratsl$Group <- factor(ratsl$Group)
ratsl$ID <- factor(ratsl$ID)

#We stardardize the ratsl dataset
ratsl <- ratsl %>% group_by(Time) %>% mutate(stweight = (Weight - mean(Weight)) / sd(Weight)) %>% ungroup()
str(ratsl)

#We plot the information from the dataset
ggplot(ratsl, aes(x = Time, y = Weight, linetype = ID)) + geom_line() + facet_grid(. ~ Group, labeller = label_both) + scale_linetype_manual(values = rep(1:6, times=4)) + theme(legend.position = "none") +   scale_y_continuous(limits = c(min(ratsl$Weight), max(ratsl$Weight)))
```
  
We can perceive an increase of weight, with certain differences, in all of the three groups.

```{r}
n <- ratsl$Time %>% unique() %>% length()

# Summarize mean and standard error
ratssum <- ratsl %>% group_by(Group, Time) %>% summarise(mean = mean(Weight), se = ( sd(Weight) / sqrt(n) )) %>% ungroup()
glimpse(ratssum)
par(mfrow = c(1,2))

#We focus first on the means
ggplot(ratssum, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

#We add the error
ggplot(ratssum, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.9,0.5)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

#We draw the boxplot
ggplot(ratsl, aes(x = factor(Time), y = Weight, fill = Group)) + geom_boxplot()
```

Boxplots with non-summarized data.

```{r}
#We create a summary for the first 8 weeks
rat8s <- ratsl %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
glimpse(rat8s)
ggplot(rat8s, aes(x=Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "pink")
```
  
  We can observe some values that can distort our results. It is best practice to neutralize them.

```{r}
#We get rid of the outlier
glimpse(rat8s)
rat8s1 <- filter(rat8s, Group == 2 & mean < 550 | Group == 1 & mean > 260 | Group == 3 & mean > 500)
glimpse(rat8s1)
ggplot(rat8s1, aes(x=Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "pink")
```
  
Once the outlier disappears, the results seem more focus around a certain point and with less dispertion of the values.  

```{r}
summary(aov(mean ~ Group, data = rat8s1))
kruskal.test(mean ~ Group, data = rat8s1)
```

We can observe that there exists statistically relevant differences between groups.
  
```{r}
#We realize a covariance analysis 
rats <- read.csv(file = "~/IODS-project/data/rats.csv", row.names = 1)
rat8s2 <- rat8s %>% mutate(baseline = rats$WD1)
glimpse(rat8s2)
fit <- lm(mean ~ baseline + Group, data = rat8s2)
summary(fit)
```

Thanks to the covariance analysis, we can see that once the original values from the first week are controlled, there exist basically no differences between the groups.

## BPRS

```{r}
#We read the data previously modified
bprsl <- read.csv(file = "~/IODS-project/data/bprsl.csv", row.names = 1)

#We convert the categorical variables into factors again
bprsl$treatment <- factor(bprsl$treatment)
bprsl$subject <- factor(bprsl$subject)
glimpse(bprsl)
```

```{r}
#We draw some initial plots to analyze them
ggplot(bprsl, aes(x = weeks, y = bprs, group = subject)) + geom_line() + facet_grid(. ~ treatment, labeller = label_both) 
```
  
There is a clear tendency of decreasing values over time in both treatment groups.


```{r}
bps_reg <- lm(bprs ~ weeks + treatment, data=bprsl)
summary(bps_reg)
```

```{r}
library(lme4)
bps_ref <- lmer(bprs ~ weeks + treatment + (1|subject), data = bprsl, REML=FALSE)
summary(bps_ref)
```

```{r}
bps_ref2 <- lmer(bprs ~ weeks + treatment + (weeks|subject), data = bprsl, REML=FALSE)
summary(bps_ref2)
anova(bps_ref, bps_ref2)
```
We see significant results in model 2.
```{r}
bps_ref3 <- lmer(bprs ~ weeks * treatment + (weeks|subject), data = bprsl, REML=FALSE)
summary(bps_ref3)
anova(bps_ref2, bps_ref3)
```
We cannot find statistically significant results for this model.
```{r}
#Create a vector of the fitted values
Fitted <- fitted(bps_ref2)
bprsl <- bprsl %>% mutate(Fitted)

#We draw a plot with the observed values
ggplot(bprsl, aes(x = weeks, y = bprs, group = subject)) + geom_line() + facet_grid(. ~ treatment, labeller = label_both) + scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +  ylim(0,100) + theme(legend.position = "top") + ggtitle("Observed values")

#Now a plot with fitted values
ggplot(bprsl, aes(x = weeks, y = Fitted, group = subject)) +
  geom_line() + facet_grid(. ~ treatment, labeller = label_both) +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +
  ylim(0, 100) +
  theme(legend.position = "top") + ggtitle("Fitted values")
```