# Regression and Model Validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

### Reading the data  
```{r}
lrn14 <- read.table("data/learning2014.txt", header = TRUE, sep = "\t")

str(lrn14)  
dim(lrn14)
```
As we can see, we have here a data frame imported as lrn14. The data reflects the relationship between learning approaches and the students achievements. This dataset includes 166 observations with 7 variables each.  The variables are gender, age, attitude, average of the deep approach, average of the strategic approach, average of the surface approachs and the points obtained in the exam.  

### Exploring the data
```{r}
pairs(lrn14[-1])
summary(lrn14)
```
In the first representation we used the *pairs()* function to show the distribution of the different variables as a function of the others.

```{r}
reg_model <- lm(points ~ attitude + stra + surf, data = lrn14)
summary(reg_model)
```

We can observe that in this multiple regression analysis we obtain a p-value that is very close to 0. This means that in this model, our explanatory variables are very likely to explain the dependent variable.  
The R-Squared represents the proportion of the variance of the points that is explained by the explanatory variables. In this case, our model explains approximately a 20% of the variance of the points.  

```{r}
par(mfrow = c(2,2))
plot(reg_model, which = c(1,2,5))
```
  
We can observe from some of the plots that the model represents a good fit. As an example, in the QQ-Plot, the great majority of the values fit very close to the line.  
On the other hand, in the plot between the Residuals and the Fitted values, we can't observe a major disperstion of the results as the fitted values increased.